---
title: "chunlei2_hw01"
author: "Chunlei Liu"
date: "8/30/2018"
output: html_document
---
Q1:

Want to estimate $\mu = E_\pi(sin(X^4)) = \int sin(X^4)*1dx$.

Because $X \sim U[0, 1]$, 1 is the probability density function of X.

Draw iid samples $x^{(1)}, x^{(2)}, ..., x^{(1000)}$ from $U[0,1]$ distribution. Estimate $\mu$ by $\hat{\mu_{n}} = \frac{1}{1000}*[sin({x^{(1)}}^4)+ ... + sin({x^{(1000)}}^4)]$

The estimate is 0.18 and the standard error is 0.24.
```{r}
#Q1
##generate 1000 samples from uniform distribution
set.seed(9301)
X1 = runif(1000, 0, 1)
mean(sin(X1^4))
sqrt(1/1000*var(sin(X1^4)))
```
Q2:

The pdf is $f(x) = \frac{4}{3}(\frac{x}{3})^3e^{-(\frac{x}{3})^4}, 0 < x < \infty$.

Then the cdf is $F(x) = \int_{0}^{x}\frac{4}{3}(\frac{a}{3})^3e^{-(\frac{a}{3})^4}da\\ \;\;\;\;\;\;\;\;= -\int_{0}^{x}e^{-(\frac{a}{3})^4}d-(\frac{a}{3})^4\\\;\;\;\;\;\;\;\;= -e^{-(\frac{a}{3})^4}|_{0}^{x}\\\;\;\;\;\;\;\;\;= -(e^{-(\frac{x}{3})^4} - 1) = U, U \sim U[0,1]$.

Because the cdf in range(0,1), so the uniform distribution can be used to simulate.
$U = 1-e^{-(\frac{x}{3})^4}$, after calculation we can get $(\frac{x}{3})^4 = -log(1-U)$. $\because x > 0, \therefore x = 3*\sqrt[4]{-log(1-U)}$.

Draw iid samples $u^{(1)}, u^{(2)}, ..., u^{(1000)}$ from $U[0,1]$

Insert them into the formula, achieve the 1000 samples from the Weibull distribution.

The mean is 2.71. The variance is 0.58.
```{r}
#Q2
##generate 1000 samples from uniform distribution
set.seed(9301)
U2 = runif(1000, 0, 1)
##calculate the inversed cdf
X2 = 3*(-log(1-U2))^(1/4)
mean(X2)
var(X2)
```
Q3:

The pdf is $f(x) = \begin{cases} 1-x, 0 \leqq x \leqq 1\\ x-1, 1< x \leqq2\end{cases}$

The cdf is $F(x) = \begin{cases} \int_{0}^{x}(1-a)da, 0 \leqq x \leqq 1\\ \int_{0}^{x}(a-1)da, 1< x \leqq2\end{cases}\\\;\;\;\;\;\;\;\; =\begin{cases} x-\frac{1}{2}x^2, 0 \leqq x \leqq 1\\ \frac{1}{2}x^2-x+1, 1< x \leqq2\end{cases}$

The area of two parts is same. So we simulate 5 samples from the uniform distribution. If the value is bigger than 0.5, it should be in the second part. Otherwise it should be in the first part.

For the first part:
$x-\frac{1}{2}x^2 = u, (0 \leqq u \leqq\frac{1}{2})\\(x-1)^2 = 1-2u\\\because 0 \leqq x \leqq 1, \therefore x =1 -\sqrt{1-2u}$

For the second part:
$1+\frac{1}{2}x^2-x = u, (\frac{1}{2} < u \leqq 1)\\(x-1)^2 = 2u-1\\\because 1 < x \leqq 2, \therefore x = 1 + \sqrt{2u-1}$

Draw iid samples $u^{(1)}, u^{(2)}, ..., u^{(5)}$ from $U[0,1]$

Divide them based on the cutoff 0.5. Insert them into the formula, achieve 5 samples from the above distribution.

```{r}
#Q3
##generate 5 samples from uniform distribution
set.seed(9301)
U3 = runif(5, 0, 1)
##if u2 > 0.5, it should belong to the second part of the function.
U3_1 = U3[U3 < 0.5]
U3_2 = U3[U3 > 0.5]
##calculate the truncated CDF
X3_1 = 1 - sqrt(1 - 2*U3_1)
X3_2 = 1 + sqrt(2*U3_2 - 1)
X3 = c(X3_1, X3_2)
X3
```


