---
title: "hw07"
author: "Chunlei Liu"
date: "10/30/2018"
output: pdf_document
---

Q1:

The target distribution is $\pi(x) \propto exp(\mu\sum_{i=1}^{d-1}x_ix_{i+1})=g(x)$, 

at time step t, $x^{(t)} = (x_1^{(1)}, ..., x_d^{(d)})$, 

at each direction i, draw $x_i^{(t+1)}$ from the conditional distribution $g(x_i^{(t+1)}|x_1^{(t+1)},..x_{i-1}^{(t+1)}, x_{i+1}^{(t)}, ..., x_d^{(t)}) = \frac{g(x_i^{(t+1)},x_1^{(t+1)},...,x_{i-1}^{(t+1)}, x_{i+1}^{(t)}, ..., x_d^{(t)}}{g(x_1^{(t+1)},...,x_{i-1}^{(t+1)}, x_{i+1}^{(t)}, ..., x_d^{(t)})}=C_i*exp(\mu x_{i-1}^{(t+1)}x_{i}+\mu x_{i}x_{i+1}^{(t)}), C_i$ is a constant.

Becasue $x_i$ only equals 1 or -1, $C_i*exp(\mu(x_{i-1}^{(t+1)}*1  + 1*x_{i+1}^{(t)})) + C_i*exp(\mu(x_{i-1}^{(t+1)}*-1+-1*x_{i+1}^{(t)})) = 1\\C_i = \frac{1}{exp(\mu(x_{i-1}^{(t+1)}+x_{i+1}^{(t)}))+exp(\mu(-x_{i-1}^{(t+1)}-x_{i+1}^{(t)}))}$

There are two special cases, 

if i = 1, $g(x_1^{(t+1)}|x_{[-1]}^{(t)}) = C_1*exp(\mu x_1x_2^{(t)})$, same rule as before $C_1 = \frac{1}{exp(\mu x_2^{(t)})+exp(-\mu x_2^{(t)})}$ 

if i = d, $g(x_d^{(t+1)}|x_{[-d]}^{(t+1)}) = C_d*exp(\mu x_{d-1}^{(t+1)}x_{d})$, same rule as before $C_d = \frac{1}{exp(\mu x_{d-1}^{(t+1)})+exp(-\mu x_{d-1}^{(t+1)})}$

```{r}
rm(list = ls())
set.seed(1)
d = 20
mu = 2
t = 50000
x_0 = sample(c(-1,1), d, replace = T)
x_matrix = matrix(0, t+1, d)
x_matrix[1, ] = x_0
gibbs = function(x, mu) {
  x_new = x
  for (i in 1:d) {
    if (i == 1) {
      c = 1/(exp(mu * x_new[i+1]) + exp(-mu * x_new[i+1]))
      prob = exp(mu * x_new[i+1])*c
      x_new[i] = rbinom(1,1,prob)*2-1
    }
    else if (i == d) {
      c = 1/(exp(mu * x_new[i-1]) + exp(-mu * x_new[i-1]))
      prob = exp(mu * x_new[i-1])*c
      x_new[i] = rbinom(1,1,prob)*2-1
    }
    else {
      c = 1/(exp(mu * (x_new[i-1] + x_new[i+1])) + exp(-mu * (x_new[i-1] + x_new[i+1])))
      prob = exp(mu * (x_new[i-1] + x_new[i+1]))*c
      x_new[i] = rbinom(1,1,prob)*2-1
    }
  }
  return(x_new)
}

for (j in 1:t) {
x_matrix[j+1,] = gibbs(x_matrix[j,], mu)
}

trunc_sample = tail(x_matrix, 20000)
M = apply(trunc_sample, 1, sum)
hist(M)
```

Q2:

$\omega = 0.04, \alpha = 2, \beta = 0.5, n =5$, The target distribution is $p(\mu,\tau|y_1,...,y_n) \propto (2\pi)^{-\frac{n+1}{2}}\tau^{\frac{n}{2}}e^{-\frac{\tau}{2}\sum_{i=1}^{n}(y_i-\mu)^2}\omega^{\frac{1}{2}}e^{-\frac{\omega}{2}\mu^2}\frac{\tau^{\alpha-1}}{\Gamma(\alpha)\beta^\alpha}e^{-\frac{\tau}{\beta}}$

at time step t, we draw

$\mu^{(t+1)}$ from $p(\mu|\tau^{(t)}, y_1...y_n) \propto e^{-\frac{\tau}{2}\sum_{i=1}^{n}(y_i-\mu)^2}e^{-\frac{\omega}{2}\mu^2} \propto e^{-\frac{n\tau+\omega}{2}(\mu - \frac{n\tau \bar{y}}{n\tau+\omega})^2} \sim N(\frac{n\tau \bar{y}}{n\tau+\omega}, \frac{1}{n\tau+\omega})$, 

$\tau^{(t+1)}$ from $p(\tau|\mu^{(t+1)},y_1...y_n) \propto \tau^{\frac{n}{2}+\alpha-1}e^{-(\frac{\sum(y_i-\mu)^2}{2}+\frac{1}{\beta})\tau} \sim \Gamma(\frac{n}{2} + \alpha, \frac{\sum(y_i-\mu)^2}{2}+\frac{1}{\beta})$

The mean is (2.1130266, 0.8561928)

```{r}
rm(list = ls())
set.seed(1)
y = c(1.9, 3.4, 0.3, 2.5, 2.6)
n = 5
omega = 0.04
alpha = 2
beta = 0.5
y_mean = mean(y)
t = 20000
x_0 = c(0, 1)
x_matrix = matrix(0, t+1, 2)
x_matrix[1, ] = x_0
gibbs = function(x){
  x_new = x
  x_new[1] = rnorm(1, (n*x_new[2]*y_mean)/(n*x_new[2]+omega), sqrt(1/(n*x_new[2]+omega)))
  x_new[2] = rgamma(1, shape = n/2+alpha, rate = (sum((y-x_new[1])^2)/2+1/beta))
  return(x_new)
}
for (i in 1:t) {
  x_matrix[i+1, ] = gibbs(x_matrix[i, ])
}
trunc_sample = tail(x_matrix, 10000)
M = apply(trunc_sample, 2, mean)
M
```

