---
author: "Yutong Dai(yutongd3@illinois.edu)"
date: "`r Sys.Date()`"
lecnum: "6"
instructor: "Yuguo Chen"
course: "STAT-525"
term: "2018 Fall"
output: 
  pdf_document:
    number_sections: yes
    template: "../textem/template.tex"
    keep_tex: no
fontsize: 10pt
bibliography: "../textem/book.bib"
link-citations: yes
linkcolor: "black"
csl: "../textem/ieee.csl"
header-includes:
  - \newcommand{\iid}{\overset{iid}\sim}
  - \usepackage{graphicx}
  - \usepackage{float}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 4)
knitr::opts_chunk$set(fig.pos = "H")
```

# Question 1

## Solution {-}

General setups:

* $T(y|x) = N(x, 4)$

* The time horizon is set to 50000, and  the burn-in sample size is set to 45000.

```{r, echo=TRUE, cache=TRUE}
ratio <- function(x,y){
 return(min(1, (exp(-y^4-y))/ (exp(-x^4-x)))) 
}
x_seq <- 0
acceptance <- 0
x_seq_length <- 50000
for (i in seq_len(x_seq_length)){
  x <- tail(x_seq, n=1)
  y <- rnorm(n=1, mean=x, sd=2)
  u <- runif(1)
  temp <- ifelse(ratio(x, y) >= u, y, x)
  if (ratio(x, y) >= u){
    acceptance <- acceptance + 1
  }
  x_seq <- c(x_seq, temp)
}
acceptance_rate <- acceptance / x_seq_length
x_seq_accepted <- tail(x_seq, 5000)
```

```{r, echo=TRUE}
muhat <- mean(x_seq_accepted)
temp <- sum(acf(x_seq_accepted, length(x_seq_accepted), plot = FALSE)$acf)
std <- sqrt(var(x_seq_accepted) * (2 * temp + 1) / length(x_seq_accepted))
```

* The acceptance rate of my algorithm is `r acceptance_rate`.

* The autocorrelation plot of my accepted samples are given in \autoref{acf}

* The mean of the desity function is `r muhat`.

* The standar error of the estimator is `r std`.

```{r, cache=TRUE, fig.cap="\\label{acf}The autocorrelation of the last 5000 samples."}
acf(x_seq_accepted, main = "")
```



**Remark**

From \autoref{compare}, we know that the samples are of good quality.

```{r, fig.cap="\\label{compare}The histgram of the last 5000 samples from the simulated Markov Chain."}
xs <- seq(-5,5, length.out = 1000)
hist(x_seq_accepted,main = "", freq=FALSE, ylim = c(0,1), ylab = "probability", xlab="x")
lines(xs, (exp(-xs^4-xs)/2.138695), type="l", col="red")
```


<!-- Sample by the rejection method -->
<!-- ```{r} -->
<!-- l <- function(x){ -->
<!--   return(exp(-x^4-x)) -->
<!-- } -->
<!-- x_rejection_seq <- NULL -->

<!-- for (i in seq_len(10000)){ -->
<!--   x <- rexp(1, rate=1) -->
<!--   if (runif(1) <= (l(x) / dexp(x))){ -->
<!--     x_rejection_seq <- c(x_rejection_seq, x) -->
<!--   } -->
<!-- } -->
<!-- hist(x_rejection_seq) -->
<!-- ``` -->

# Question 2

## Solution {-}


General setups:

$T(T_{t+1}|T_{t}) = 0.5\pi(T_{t}^+)I(T_{t+1}=T_{t}^+) + 0.5\pi(T_{t}^+)I(T_{t+1}=T_{t}^-)$, where $T_{t}^+, T_{t}^-$ represents the two transformation patterns respectively. Clearly, this is symmetric.

* The time horizon is set to 20000, and  the burn-in sample size is set to 19500.

```{r, echo=TRUE, cache=TRUE}
rm(list=ls())
T0 <- read.delim("xid-36891322_1.txt", row.names=1, sep="", stringsAsFactors=FALSE)[1:12, 1:12]

easy_ratio <- function(seq1, seq2){
  ratio <- 1
  seq1 <- sort(seq1); seq2 <- sort(seq2)
  for (i in seq_len(length(seq1))){
    if (seq1[i] > seq2[i]){
      ratio <- ratio * prod(seq((seq2[i] + 1),  seq1[i]))
    }else if(seq1[i] < seq2[i]){
      ratio <- ratio / prod(seq((seq1[i] + 1),  seq2[i]))
    }
  }
  return(ratio)
}

sample_new <- function(Tx){
  #browser()
  row_idx <- sort(sample.int(12, 2, replace = FALSE))
  col_idx <- sort(sample.int(12, 2, replace = FALSE))
  if (rbinom(1,1,0.5) == 1){
      ratio <- (Tx[row_idx[1],col_idx[2]] * Tx[row_idx[2],col_idx[1]]) / 
        ((Tx[row_idx[1],col_idx[1]] + 1) * (Tx[row_idx[2],col_idx[2]] + 1))
      change <- matrix(c(1,-1,-1,1), byrow = T, nrow=2)
  }else{
      ratio <- (Tx[row_idx[1],col_idx[1]] * Tx[row_idx[2],col_idx[2]]) /
        ((Tx[row_idx[1],col_idx[2]] + 1) * (Tx[row_idx[2],col_idx[1]] + 1))
      change <- matrix(c(-1,1,1,-1), byrow = T, nrow=2)
    }
  rxy <- min(1, ratio)
  if (runif(1) <= rxy){
    Ty <- Tx
    Ty[row_idx[1],col_idx[1]] <- Tx[row_idx[1],col_idx[1]] + change[1,1]
    Ty[row_idx[1],col_idx[2]] <- Tx[row_idx[1],col_idx[2]] + change[1,2]
    Ty[row_idx[2],col_idx[1]] <- Tx[row_idx[2],col_idx[1]] + change[2,1]
    Ty[row_idx[2],col_idx[2]] <- Tx[row_idx[2],col_idx[2]] + change[2,2]
    accept <- 1
  }else{
    Ty <- Tx
    accept <- 0
  }
  PTy_PT0 <- 1 / easy_ratio(as.vector(as.matrix(Ty)), as.vector(as.matrix(T0)))
  results <- list(Ty=Ty, accept=accept, PTy_PT0 = PTy_PT0)
  return(results)
}

T_seq <- list()
T_seq[[1]] <- T0
acceptance <- 0
T_seq_length <- 20000
compare_seq <- NULL
for (i in seq_len(T_seq_length)){
  Tx <- T_seq[[i]]
  results <- sample_new(Tx)
  T_seq[[i+1]] <- results$Ty
  acceptance <- acceptance + results$accept
  compare_seq <- c(compare_seq, results$PTy_PT0 <= 1)
}
acceptance_rate <- acceptance / T_seq_length
```

```{r, echo=TRUE}
seq_accepted_final <- tail(compare_seq, 500)
p_val <- mean(seq_accepted_final)
h_T <- as.numeric(seq_accepted_final)
h_T <- as.numeric(seq_accepted_final)
temp <- 1 + 2 * (sum(acf(h_T, 500, plot = FALSE)$acf) - 1)
std <- sqrt(abs(var(h_T) * temp) / 500)
```

So the

* P-value is $`r p_val`$.
* And the standar error is $`r std`$.

**Remark:**

The theoretical result is approximately

```{r}
fisher.test(T0, simulate.p.value=TRUE, B=10000)
```

